---
title: "Weight Lifting Activity Mode Prediction Model"
author: "E.C.S"
date: "Sunday, April 26, 2015"
output: html_document
---


-- Load caret 
```{r}

library(caret)

```





-- Read in traing and test data 

```{r}

setwd("~/Data Science Specialization/machine_learning/week3_3")


trainData <- read.csv('pml-training.csv',na.strings = c("NA",""))


testData <- read.csv('pml-testing.csv',na.strings = c("NA",""))

```


-- Split training set into training and test partitions
```{r}

set.seed(1234)

trainPart <- createDataPartition(trainData$classe, list=FALSE, p=.80)
trainSet <- trainData[trainPart,]
testSet <- trainData[-trainPart,]

dim(trainSet)
dim(testSet) 
```

-- Remove columns with near zero variance from prediction attributes...
```{r}
nzv <- nearZeroVar(trainSet) 
trainSet <- trainSet[-nzv]
testSet <- testSet[-nzv]
testData <- testData[-nzv]

dim(trainSet)
dim(testSet) 

```

-- Retain numeric attributes only for use in prediction model... 

```{r}

numeric_attr = which(lapply(trainSet,class) %in% c('numeric') ) 

```

-- use preProcess.knnImpute to impute missing data values... lli

```{r}

library(caret)
library(randomForest)
library(RANN)

beta <- preProcess(trainSet[,numeric_attr], method=c('knnImpute'))



trainPredict <- cbind(trainSet$classe, predict(beta, trainSet[,numeric_attr]))
testValPredict <- cbind(testSet$classe, predict(beta, testSet[,numeric_attr]))
testPredict <- predict(beta, testData[,numeric_attr])

names(trainPredict)[1] <- 'classe'
names(testValPredict)[1] <- 'classe'

```

```{r}

randfor <- randomForest(classe ~ ., trainPredict, ntree=1000, mtry=24)

```



--Use test data to evaluate quality of the prediction model..
```{r}

answers <- predict(randfor, testPredict)
answers

```

--Post answers one answer to each file in subdirectory given ...
```{r}

setwd("./Answers")

unlink("problem_id_*")

pml_write_files = function(x){
     n = length(x)
     for(i in 1:n){
         filename = paste0("problem_id_",i,".txt")
         write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
     }
 }


pml_write_files(answers)

```


--Results: 

    The model seems to have good prediction ability for the test data set provided. 


--Reference publication basis for predictive Model:

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. 
Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013. 

